---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r preproc, echo=TRUE, message=F}
library(dplyr)
library(ggplot2)
unzip(zipfile = "activity.zip")
act <- read.csv("activity.csv")
act$date <- as.Date(act$date)

```

## What is mean total number of steps taken per day?
```{r }
naAct <- act[complete.cases(act),] 
stepsByDay <- summarise(group_by(naAct,date), steps=sum(steps))
```
Another ways to calculate the same:  
* tapply(naAct\$steps, naAct$date, sum)  
* kk <-aggregate(steps ~ date, naAct, sum)  

```{r}
hist(stepsByDay$steps, breaks=15)
```
   
The steps by day mean is `r as.character(round(mean(stepsByDay$steps),0))` and its median   `r median(stepsByDay$steps)`


## What is the average daily activity pattern?
```{r echo=T }
naAct$dayInterval <- 1:(24*60/5)
stepsByInt <- summarise(group_by(naAct,dayInterval), steps=mean(steps))

p1 <- ggplot(stepsByInt, aes(x = dayInterval, y = steps))
p1 + geom_line()

```
    
This is the 5-minute interval that, on average across all the days in the dataset, contains the maximum number of steps, and that number of steps:   
```{r}
filter(stepsByInt, steps==max(steps))
```

## Imputing missing values

There are `r nrow(act)-nrow(naAct)` rows with NA values.

To fill these missing values, we are going to use the mean interval . We assume that the number of
steps in one interval is similar to the number of steps of the same interval in other days. 

```{r missingValues1, echo=TRUE}
act$dayInterval <- 1:(24*60/5)
actCo <- act
actCo <- merge(actCo,stepsByInt,by="dayInterval")

pos_na <- which(is.na(actCo$steps.x))
actCo[pos_na,"steps.x"] <-  actCo[pos_na,"steps.y"]

stepsByDay_Co <- summarise(group_by(actCo,date), steps=sum(steps.x))
hist(stepsByDay_Co$step, breaks=15)
```
   
The steps by day mean is `r as.character(round(mean(stepsByDay_Co$steps),0))` and its median   `r median(stepsByDay_Co$steps)`
```



## Are there differences in activity patterns between weekdays and weekends?
